Left-Linear Rewriting in Adhesive Categories

### 2024-07-08 concur/final.tex
typos

\noindent
\parbox{8cm}{
  \begin{definition}[abstraction equivalence]
    Let  $\X$ be an $\mathcal{M}$-adhesive category and 
    $(\X, \R)$ a left-linear rewriting system. 
    An \emph{abstraction equivalence} between derivations $\der{D}$
    and $\der{D'}$ with the same length and
    $r(\der{D})=r(\der{D}')$ is a family of isomorphisms
    $\{\phi_X\}_{X\in \{G_i,D_i\}}$ such that the diagram on the right commutes
    for $i\in [0, \lgh(\der{D})]$. We say that $\der{D}$ are $\der{D}'$ are \emph{abstraction
      equivalent}, written $\der{D}\equiv_a \der{D}'$, if there exists an abstraction equivalence between
    them.
  \end{definition}
}
\parbox{4cm}{\vspace{-1.35em}$\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i}
    \ar@{>->}[l]_{f'_i}&G'_{i+1}\\ L_i \ar[u]^{m'_i} \ar[d]_{m_i}& K_i
    \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar@{>->}[l]_{l_i}
    &R_i \ar[u]^{h'_i} \ar[d]_{h_i}\\G_i
    \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar@{>->}[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole
    \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{G_{i+1}}}}$}

@@

We  recall some %concepts and
notation on permutations. 
% \begin{definition}
A \emph{permutation} on
$\interval[0]{n}$ is a bijection
$\sigma : \interval[0]{n} \to \interval[0]{n}$. It is a
\emph{transposition} $\nu$ if there are $i, j \in \interval{n}$,
$i \neq j$ such that $\sigma(i)=j$, $\sigma(j) = i$, and
$\sigma(k) = k$ otherwise; it is denoted as $\transp{i}[j]$.
%Such a transposition is
%% normally
%denoted as
%$\transp{i}[j]$ and we use the letter $\nu$ for transpositions.
%% 
Given a permutation $\perm$, % on $\interval[0]{n}$
an \emph{inversion}
for $\sigma$ is a pair $(i,j)$ such that $i<j$ and
$\sigma(j)< \sigma(i)$; $\inv{\sigma}$ denotes the set of
inversions of $\perm$.
%\end{definition}

@@

\noindent 
\parbox{11.5cm}{\begin{proposition}\label{prop:pbpo} Let $\mathcal{A}$ be a class of arrows stable under pushouts and containing all the isomorphisms.  If $m\colon A\to C$ is mono and belongs to $\mathcal{A}$, then every $\mathcal{A}$-Van Kampen square
	is also a pullback square and $n$ is a monomorphism.
\end{proposition}}
\parbox{2cm}{\vspace{1.5ex}$\xymatrix{A\ar[r]^{g} \ar[d]_{m} & B \ar[d]^{n} \\ C \ar[r]_{f}  & D}$}

@@

	The thesis now follows immediately from \Cref{lem:varie}. \qedhere
\end{proof}
}
\noindent
\parbox{10cm}{
	\begin{corollary}[Uniqueness of pushout complements]\label{lem:pocomp}
		Let $\X$ be a $\mathcal{M}$-adhesive category. Given $m\colon X\mto Y$ in $\mathcal{M}$ and $f\colon Y\to Z$, let $(h_1, k_1)$ and $(h_2, k_2)$ be respectively the pushout complements of $m$  and $f$ depicted aside. Then there exists a unique isomorphism $\phi\colon W_1\to W_2$ making the diagram on the right commutative.
	\end{corollary}}
\parbox{3cm}{\vspace{1em}$\xymatrix{&X \ar@{>->}[r]^{m} \ar[d]_{h_1} \ar@/_.3cm/[dl]_{h_2}& Y \ar[d]^{f} \\ W_2 \ar@{>->}@/_.4cm/[rr]_{k_2} &W_1 \ar@{.>}[l]_{\phi} \ar@{>->}[r]^{k_1} & Z }$}

>>>>

\noindent
\parbox{8cm}{
  \begin{definition}[abstraction equivalence]
    Let  $\X$ be an $\mathcal{M}$-adhesive category and 
    $(\X, \R)$ a left-linear rewriting system. 
    An \emph{abstraction equivalence} between derivations $\der{D}$
    and $\der{D'}$ with the same length and
    $r(\der{D})=r(\der{D}')$ is a family of isomorphisms
    $\{\phi_X\}_{X\in \{G_i,D_i\}}$ such that the diagram on the right commutes
    for $i\in [0, \lgh(\der{D}))$. We say that $\der{D}$ are $\der{D}'$ are \emph{abstraction
      equivalent}, written $\der{D}\equiv_a \der{D}'$, if there exists an abstraction equivalence between
    them.
  \end{definition}
}
\parbox{4cm}{\vspace{-1.35em}$\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i}
    \ar@{>->}[l]_{f'_i}&G'_{i+1}\\ L_i \ar[u]^{m'_i} \ar[d]_{m_i}& K_i
    \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar@{>->}[l]_{l_i}
    &R_i \ar[u]^{h'_i} \ar[d]_{h_i}\\G_i
    \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar@{>->}[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole
    \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{G_{i+1}}}}$}

@@

We  recall some %concepts and
notations on permutations. 
% \begin{definition}
A \emph{permutation} on
$\interval[0]{n}$ is a bijection
$\sigma : \interval[0]{n} \to \interval[0]{n}$. It is a
\emph{transposition} $\nu$ if there are $i, j \in \interval{n}$,
$i \neq j$ such that $\sigma(i)=j$, $\sigma(j) = i$, and
$\sigma(k) = k$ otherwise; it is denoted as $\transp{i}[j]$.
%Such a transposition is
%% normally
%denoted as
%$\transp{i}[j]$ and we use the letter $\nu$ for transpositions.
%% 
Given a permutation $\perm$, % on $\interval[0]{n}$
an \emph{inversion}
for $\sigma$ is a pair $(i,j)$ such that $i<j$ and
$\sigma(j)< \sigma(i)$; $\inv{\sigma}$ denotes the set of
inversions of $\perm$.
%\end{definition}

@@

\noindent 
\parbox{11.5cm}{\begin{proposition}\label{prop:pbpo} Let $\mathcal{A}$ be a class of arrows stable under pushouts and containing all the isomorphisms.  If the square to the right is $\mathcal{A}$-Van Kampen and $m\colon A\to C$ is mono and belongs to $\mathcal{A}$, then the square is a pullback and $n$ is a monomorphism.
\end{proposition}}
\parbox{2cm}{\vspace{1.5ex}$\xymatrix{A\ar[r]^{g} \ar[d]_{m} & B \ar[d]^{n} \\ C \ar[r]_{f}  & D}$}

@@

	The thesis now follows immediately from \Cref{lem:varie}. \qedhere
\end{proof}
}
\noindent
\parbox{10cm}{
	\begin{corollary}[Uniqueness of pushout complements]\label{lem:pocomp}
		Let $\X$ be a $\mathcal{M}$-adhesive category. Given $m\colon X\mto Y$ in $\mathcal{M}$ and $f\colon Y\to Z$, let $(h_1, k_1)$ and $(h_2, k_2)$ be pushout complements of $m$  and $f$ depicted aside. Then there exists a unique isomorphism $\phi\colon W_1\to W_2$ making the diagram on the right commutative.
	\end{corollary}}
\parbox{3cm}{\vspace{1em}$\xymatrix{&X \ar@{>->}[r]^{m} \ar[d]_{h_1} \ar@/_.3cm/[dl]_{h_2}& Y \ar[d]^{f} \\ W_2 \ar@{>->}@/_.4cm/[rr]_{k_2} &W_1 \ar@{.>}[l]_{\phi} \ar@{>->}[r]^{k_1} & Z }$}
